---
title: "Living in a bottle: Bacteria from sediment-associated Mediterranean waste and potential growth on polyethylene terephthalate (PET)"
author: "Adriel Latorre-Pérez - Darwin Bioprospecting Excellence S.L."
date: "2020-2021"
output: html_document
---

```{r}
# Remove R data
rm (list = ls())
```


# Data import

Data has been already imported into phyloseq, so the .Rdata file has to be loaded:

```{r}
library(phyloseq)
load("./study.RData")
```

physeq1 = phyloseq object at the ASV level (absolute abundances)

Check the number of seqs per sample and the overall sequencing stats.

```{r}
sample_sums(physeq1)
summary(sample_sums(physeq1))
```

# Rarefaction curves

```{r}
library("iNEXT")
library(ggplot2)
Rarefaction <- iNEXT(as.data.frame(otu_table(physeq1)), q=0, datatype="abundance")
Rareplot1 <- ggiNEXT(x=Rarefaction, type=1) + theme_bw() +
  geom_line(size=0.1) +
  geom_point(size=0, na.rm = TRUE) +
  scale_shape_manual(values=c(rep(20,55))) + xlab("Number of sequences") + ylab("Richness (Observed OTUs)")
Rareplot1
```

# Alpha diversity

By each sample type:

```{r}
library(ggplot2)
p = plot_richness(rarefy_even_depth(physeq1, rngseed = 711), x="SampleType2", color="SampleType2", measures=c("Observed","Simpson", "Shannon")) + xlab("") + ylab("") + geom_point() + geom_boxplot(alpha = .4) + theme_light()
p
```

**Find differences between control and plastic**

*Richness:*

```{r echo=FALSE, warning=FALSE, comment=FALSE}
rich = estimate_richness(physeq1)
wilcox.test(rich$Observed[sample_data(physeq1)$SampleType2 == "Control"], rich$Observed[sample_data(physeq1)$SampleType2 == "Plastic"])
```

NO!

*Simpson*

```{r echo=FALSE, warning=FALSE, comment=FALSE}
rich = estimate_richness(physeq1)
wilcox.test(rich$Simpson[sample_data(physeq1)$SampleType2 == "Control"], rich$Simpson[sample_data(physeq1)$SampleType2 == "Plastic"])
```

No

*Shannon*

```{r echo=FALSE, warning=FALSE, comment=FALSE}
rich = estimate_richness(physeq1)
wilcox.test(rich$Shannon[sample_data(physeq1)$SampleType2 == "Control"], rich$Shannon[sample_data(physeq1)$SampleType2 == "Plastic"])
```

No.

**Control vs. can**

*Richness:*

```{r echo=FALSE, warning=FALSE, comment=FALSE}
rich = estimate_richness(physeq1)
wilcox.test(rich$Observed[sample_data(physeq1)$SampleType2 == "Control"], rich$Observed[sample_data(physeq1)$SampleType2 == "Can"])
```

NO

*Simpson*

```{r echo=FALSE, warning=FALSE, comment=FALSE}
rich = estimate_richness(physeq1)
wilcox.test(rich$Simpson[sample_data(physeq1)$SampleType2 == "Control"], rich$Simpson[sample_data(physeq1)$SampleType2 == "Can"])
```

NO

*Shannon*

```{r echo=FALSE, warning=FALSE, comment=FALSE}
rich = estimate_richness(physeq1)
wilcox.test(rich$Shannon[sample_data(physeq1)$SampleType2 == "Control"], rich$Shannon[sample_data(physeq1)$SampleType2 == "Can"])
```

NO

**Can vs. plastic**

*Richness:*

```{r echo=FALSE, warning=FALSE, comment=FALSE}
rich = estimate_richness(physeq1)
wilcox.test(rich$Observed[sample_data(physeq1)$SampleType2 == "Plastic"], rich$Observed[sample_data(physeq1)$SampleType2 == "Can"])
```

NO

*Simpson*

```{r echo=FALSE, warning=FALSE, comment=FALSE}
rich = estimate_richness(physeq1)
wilcox.test(rich$Simpson[sample_data(physeq1)$SampleType2 == "Plastic"], rich$Simpson[sample_data(physeq1)$SampleType2 == "Can"])
```

NO

*Shannon*

```{r echo=FALSE, warning=FALSE, comment=FALSE}
rich = estimate_richness(physeq1)
wilcox.test(rich$Shannon[sample_data(physeq1)$SampleType2 == "Plastic"], rich$Shannon[sample_data(physeq1)$SampleType2 == "Can"])
```

No


**CONCLUSION: no significant differences in terms of alpha diversity**

--------------------------------------------------------------------------------------------


# Principal Coordinates Analysis (PCoA)

## ASV level

Calculate relative abudnances

```{r}
physeq1_rel  = transform_sample_counts(physeq1, function(x) x / sum(x)*100 )
sample_sums(physeq1_rel)
```

And create the graph

```{r}
library(ggrepel)
library(ggplot2)
# Calculate distance matrix
brayDist <- phyloseq::distance(physeq1_rel, method="bray")
# Calculate ordination
iMDS  <- ordinate(physeq1_rel, distance=brayDist, method = "PCoA")
## Make plot
# Create plot, store as temp variable, p
p <- plot_ordination(physeq1_rel, iMDS, color ="SampleType2") + theme_light()
# Costumize the plot
p = p + geom_point(aes(size=3, alpha = 0.6)) + geom_text_repel(aes(label = SampleName))
p
```


## PERMANOVA

```{r}
library(vegan)
physeq33 <- as(sample_data(physeq1_rel), "data.frame")
adonis2(brayDist ~ SampleType2, data = physeq33)
```

No significant differences

## Genus level

Calculate relative abundances

```{r}
physeq_R6 <- tax_glom(physeq1, taxrank = rank_names(physeq1)[6], NArm = FALSE)
physeq_R6_rel  = transform_sample_counts(physeq_R6, function(x) x / sum(x)*100 )
sample_sums(physeq_R6_rel)
```

## PCoA genus level

```{r}
library(ggrepel)
library(ggplot2)
# Calculate distance matrix
brayDist <- phyloseq::distance(physeq_R6_rel, method="bray")
# Calculate ordination
iMDS  <- ordinate(physeq_R6_rel, distance=brayDist, method = "PCoA")
## Make plot
# Create plot, store as temp variable, p
p <- plot_ordination(physeq_R6_rel, iMDS, color ="SampleType2") + theme_light()
# Costumize the plot
p = p + geom_point(aes(size=3, alpha = 0.6)) + geom_text_repel(aes(label = SampleName))
p
ggsave("./Figures/PCoA-L6-all.svg", plot = p)
```

## PERMANOVA

```{r}
library(vegan)
physeq33 <- as(sample_data(physeq_R6_rel), "data.frame")
adonis2(brayDist ~ SampleType2, data = physeq33)
```

# PCoA without Plastic-2

Let's remove this sample, which is clearly different to the rest by whatever reason.

## ASV level

```{r}
physeq1_noP2 = subset_samples(physeq1, SampleName != "Plastic-2")
physeq1_noP2
```

```{r}
physeq1_noP2_rel  = transform_sample_counts(physeq1_noP2, function(x) x / sum(x)*100 )
sample_sums(physeq1_noP2_rel)
```

```{r}
library(ggrepel)
library(ggplot2)
# Calculate distance matrix
brayDist <- phyloseq::distance(physeq1_noP2_rel, method="bray")
# Calculate ordination
iMDS  <- ordinate(physeq1_noP2_rel, distance=brayDist, method = "PCoA")
## Make plot
# Create plot, store as temp variable, p
p <- plot_ordination(physeq1_noP2_rel, iMDS, color ="SampleType2") + theme_light()
# Costumize the plot
p = p + geom_point(aes(size=3, alpha = 0.6)) + geom_text_repel(aes(label = SampleName))
p
# ggsave("./Figures/PCoA-OTU-noP2.png", plot = p, device = "png")
```

```{r}
library(vegan)
physeq33 <- as(sample_data(physeq1_noP2_rel), "data.frame")
adonis2(brayDist ~ SampleType2, data = physeq33)
```

## Genus level

```{r}
physeq_R6_noP2 = subset_samples(physeq_R6, SampleName != "Plastic-2")
physeq_R6_noP2
```

```{r}
physeq_R6_noP2_rel  = transform_sample_counts(physeq_R6_noP2, function(x) x / sum(x)*100 )
sample_sums(physeq_R6_noP2_rel)
```

```{r}
library(ggrepel)
library(ggplot2)
# Calculate distance matrix
brayDist <- phyloseq::distance(physeq_R6_noP2_rel, method="bray")
# Calculate ordination
iMDS  <- ordinate(physeq_R6_noP2_rel, distance=brayDist, method = "PCoA")
## Make plot
# Create plot, store as temp variable, p
p <- plot_ordination(physeq_R6_noP2_rel, iMDS, color ="SampleType2") + theme_light()
# Costumize the plot
p = p + geom_point(aes(size=3, alpha = 0.6)) + geom_text_repel(aes(label = SampleName))
p
ggsave("./Figures/PCoA-L6-noP2.svg", plot = p)
```

```{r}
library(vegan)
physeq33 <- as(sample_data(physeq_R6_noP2_rel), "data.frame")
adonis2(brayDist ~ SampleType2, data = physeq33)
```

# PCoA Can vs. Plastic

Only Can and Plastic samples (without Plastic-2)

## ASV level

```{r}
physeq1_set = subset_samples(physeq1, SampleType2 == "Can" | SampleType2 == "Plastic")
physeq1_set = subset_samples(physeq1_set, SampleName != "Plastic-2")
sample_sums(physeq1_set)
```

```{r}
physeq1_set_rel  = transform_sample_counts(physeq1_set, function(x) x / sum(x)*100 )
sample_sums(physeq1_set_rel)
```

```{r}
library(ggrepel)
library(ggplot2)
# Calculate distance matrix
brayDist <- phyloseq::distance(physeq1_set_rel, method="bray")
# Calculate ordination
iMDS  <- ordinate(physeq1_set_rel, distance=brayDist, method = "PCoA")
## Make plot
# Create plot, store as temp variable, p
p <- plot_ordination(physeq1_set_rel, iMDS, color ="SampleType2") + theme_light()
# Costumize the plot
p = p + geom_point(aes(size=3, alpha = 0.6)) + geom_text_repel(aes(label = SampleName))
p
# ggsave("./Figures/PCoA-OTU-noP2.png", plot = p, device = "png")
```

```{r}
library(vegan)
physeq33 <- as(sample_data(physeq1_set_rel), "data.frame")
adonis2(brayDist ~ SampleType2, data = physeq33)
```

No differences

## GEnus level

```{r}
physeq_R6_set = subset_samples(physeq_R6, SampleType2 == "Can" | SampleType2 == "Plastic")
physeq_R6_set = subset_samples(physeq_R6_set, SampleName != "Plastic-2")
physeq_R6_set
```

```{r}
physeq_R6_set_rel  = transform_sample_counts(physeq_R6_set, function(x) x / sum(x)*100 )
sample_sums(physeq_R6_set_rel)
```

```{r}
library(ggrepel)
library(ggplot2)
# Calculate distance matrix
brayDist <- phyloseq::distance(physeq_R6_set_rel, method="bray")
# Calculate ordination
iMDS  <- ordinate(physeq_R6_set_rel, distance=brayDist, method = "PCoA")
## Make plot
# Create plot, store as temp variable, p
p <- plot_ordination(physeq_R6_set_rel, iMDS, color ="SampleType2") + theme_light()
# Costumize the plot
p = p + geom_point(aes(size=3, alpha = 0.6)) + geom_text_repel(aes(label = SampleName))
p
ggsave("./Figures/PCoA-L6-can&plastic.svg", plot = p)
```

```{r}
library(vegan)
physeq33 <- as(sample_data(physeq_R6_set_rel), "data.frame")
adonis2(brayDist ~ SampleType2, data = physeq33)
```

No differences

---------------------------------------------------------------------

# Bacterial taxonomic profiles

# Genus level

```{r echo=FALSE, reuslts=FALSE, warning=FALSE, comment=FALSE}
# Tabla nivel de genero para trabajar con EXCEL
all = phyloseq_to_df(physeq_R6_rel, sorting = NULL)
write.table(all, row.names = FALSE, file = "./rel_otu_table_L6.csv", quote = FALSE, sep = '\t', dec = '.')
```

I've modified this table manually on Excel to select the top 20 genera. Let's load this new table

```{r echo=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
top20 = read.csv2("./Taxonomy-tables/top_L6_noOther.csv", sep = ",", dec = ".")
colnames(top20)[c(6, 11, 16)] = c(" ", "  ", "   ")
top20
```

```{r echo=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
library(tidyr)
# The arguments to gather():
# - data: Data object
# - key: Name of new key column (made from names of data columns)
# - value: Name of new value column
# - ...: Names of source columns that contain values
# - factor_key: Treat the new key column as a factor (instead of character vector)
long.top20 <- gather(top20, Condition, Relative_abundance, Control.1:Biofilm.1, factor_key=TRUE)
long.top20
```

Do the graph

```{r echo=FALSE, comment=FALSE}
library(ggplot2)
library(forcats)
library(RColorBrewer)

# Creo una paleta semi-automática
getPalette = colorRampPalette(brewer.pal(9, "Set1"))
palette = rev(getPalette(20))
palette[c(1,20)] = c("#d9d9d9", "#ff8282")

p = ggplot(long.top20, aes(fill = fct_reorder(Taxon, Order, .desc = TRUE), y=Relative_abundance, x=Condition)) + 
    geom_bar( stat="identity", color = "black")
p = p + ylab("Relative abundance (%)") + xlab("") + guides(fill=guide_legend(ncol=2, title = "Taxa"))
p = p + theme(axis.text.x = element_text(angle = 45, hjust = 1), panel.background = element_rect(fill = "white", colour = "grey"))
p = p + scale_fill_manual(values = palette)
p
# ggsave(filename = "./Figures/genera.png", plot = p, device = "png")
```

Make the graph interactive

```{r}
library(ggplot2)
library(plotly)
fig = ggplotly(p)

htmlwidgets::saveWidget(fig, "genera.html")
```

## Phylum level taxonomy

Just repeat the commands above but changing some parameters.

```{r}
physeq_R2 <- tax_glom(physeq1, taxrank = rank_names(physeq1)[2], NArm = FALSE)
physeqfull_relabund_R2  = transform_sample_counts(physeq_R2, function(x) x / sum(x)*100)
physeq_R2
```


```{r echo=FALSE, reuslts=FALSE, warning=FALSE, comment=FALSE}
# Tabla nivel de genero para trabajar con EXCEL
all = phyloseq_to_df(physeqfull_relabund_R2, sorting = NULL)
write.table(all, row.names = FALSE, file = "./rel_otu_table_L2.csv", quote = FALSE, sep = '\t', dec = '.')
```

Modify the table in Excel and load it again.

```{r}
top20 = read.csv2("./Taxonomy-tables/top_L2.csv", sep = ",", dec = ".")
colnames(top20)[c(6, 11, 16)] = c(" ", "  ", "   ")
top20
```

Change to long format.

```{r}
library(tidyr)
# The arguments to gather():
# - data: Data object
# - key: Name of new key column (made from names of data columns)
# - value: Name of new value column
# - ...: Names of source columns that contain values
# - factor_key: Treat the new key column as a factor (instead of character vector)
long.top20 <- gather(top20, Condition, Relative_abundance, Control.1:Biofilm.1, factor_key=TRUE)
long.top20
```

Create the figure

```{r}
library(forcats)
library(ggplot2)
library(RColorBrewer)

# Vamos a construir una paleta de colores semi-automática con 20 colores
getPalette = colorRampPalette(brewer.pal(9, "Set3"))

p = ggplot(long.top20, aes(fill = fct_reorder(Taxon, Order, .desc = TRUE), y=Relative_abundance, x=Condition)) + 
    geom_bar( stat="identity", color = "black")
p = p + ylab("Relative abundance (%)") + xlab("") + guides(fill=guide_legend(ncol=2, title = "Taxa"))
p = p + theme(axis.text.x = element_text(angle = 45, hjust = 1), panel.background = element_rect(fill = "white", colour = "grey"))
p = p + scale_fill_manual(values = rev(getPalette(20))) 
p
# ggsave(filename = "./Figures/Phylum.png", plot = p, device = "png")
```

Make it interactive

```{r}
library(ggplot2)
library(plotly)
fig = ggplotly(p)

htmlwidgets::saveWidget(fig, "phylum.html")
```

---------------------------------------------------------------------

# Differential abundance analysis with DESeq2 - Genus Level

We are going to do pairwise test for the different sample types.

## Can Vs Control

```{r results=FALSE, warning=FALSE, echo = FALSE, message=FALSE}
library(DESeq2)
set1 = subset_samples(physeq_R6, SampleType2 == "Can" | SampleType2 == "Control")
sample_data(set1)$SampleType2 = relevel(sample_data(set1)$SampleType2, "Control")
# TRASFORMATION TO DESeq2
diagdds = phyloseq_to_deseq2(set1, ~ SampleType2)
diagdds = DESeq(diagdds, test="Wald", fitType="parametric")
```

Sort by p-value, remove NA values and format the results.

```{r}
res = results(diagdds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(set1)[rownames(sigtab), ], "matrix"))
sigtab
write.table(sigtab, "./Deseq2-Tables/CanVsControl.csv", sep = ",",row.names = FALSE)
```


## Plastic Vs Control

Just repeat the same commands

```{r results=FALSE, warning=FALSE, echo = FALSE, message=FALSE}
library(DESeq2)
set1 = subset_samples(physeq_R6, SampleType2 == "Plastic" | SampleType2 == "Control")
sample_data(set1)$SampleType2 = relevel(sample_data(set1)$SampleType2, "Control")
# TRASFORMATION TO DESeq2
diagdds = phyloseq_to_deseq2(set1, ~ SampleType2)
diagdds = DESeq(diagdds, test="Wald", fitType="parametric")
```


```{r}
res = results(diagdds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(set1)[rownames(sigtab), ], "matrix"))
sigtab
write.table(sigtab, "./Deseq2-Tables/PlasticVsControl.csv", sep = ",",row.names = FALSE)
```

## Plastic Vs Can

```{r results=FALSE, warning=FALSE, echo = FALSE, message=FALSE}
library(DESeq2)
set1 = subset_samples(physeq_R6, SampleType2 == "Plastic" | SampleType2 == "Can")
sample_data(set1)$SampleType2 = relevel(sample_data(set1)$SampleType2, "Can")
# TRASFORMATION TO DESeq2
diagdds = phyloseq_to_deseq2(set1, ~ SampleType2)
diagdds = DESeq(diagdds, test="Wald", fitType="parametric")
```


```{r}
res = results(diagdds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(set1)[rownames(sigtab), ], "matrix"))
sigtab
write.table(sigtab, "./Deseq2-Tables/PlasticVsCan.csv", sep = ",",row.names = FALSE)
```

# Differential abundance analysis with DESeq2 - Phylum Level

## Can Vs Control

```{r results=FALSE, warning=FALSE, echo = FALSE, message=FALSE}
library(DESeq2)
set1 = subset_samples(physeq_R2, SampleType2 == "Can" | SampleType2 == "Control")
sample_data(set1)$SampleType2 = relevel(sample_data(set1)$SampleType2, "Control")
# TRASFORMATION TO DESeq2
diagdds = phyloseq_to_deseq2(set1, ~ SampleType2)
diagdds = DESeq(diagdds, test="Wald", fitType="parametric")
```

```{r}
res = results(diagdds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(set1)[rownames(sigtab), ], "matrix"))
sigtab
write.table(sigtab, "./Deseq2-Tables/CanVsControl-phylum.csv", sep = ",",row.names = FALSE)
```


## Plastic Vs Control

```{r results=FALSE, warning=FALSE, echo = FALSE, message=FALSE}
library(DESeq2)
set1 = subset_samples(physeq_R2, SampleType2 == "Plastic" | SampleType2 == "Control")
sample_data(set1)$SampleType2 = relevel(sample_data(set1)$SampleType2, "Control")
# TRASFORMATION TO DESeq2
diagdds = phyloseq_to_deseq2(set1, ~ SampleType2)
diagdds = DESeq(diagdds, test="Wald", fitType="parametric")
```


```{r}
res = results(diagdds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(set1)[rownames(sigtab), ], "matrix"))
sigtab
write.table(sigtab, "./Deseq2-Tables/PlasticVsControl-phylum.csv", sep = ",",row.names = FALSE)
```

## Plastic Vs Can

```{r results=FALSE, warning=FALSE, echo = FALSE, message=FALSE}
library(DESeq2)
set1 = subset_samples(physeq_R2, SampleType2 == "Plastic" | SampleType2 == "Can")
sample_data(set1)$SampleType2 = relevel(sample_data(set1)$SampleType2, "Can")
# TRASFORMATION TO DESeq2
diagdds = phyloseq_to_deseq2(set1, ~ SampleType2)
diagdds = DESeq(diagdds, test="Wald", fitType="parametric")
```


```{r}
res = results(diagdds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(set1)[rownames(sigtab), ], "matrix"))
sigtab
write.table(sigtab, "./Deseq2-Tables/PlasticVsCan_phylum.csv", sep = ",",row.names = FALSE)
```



--------------------------------------------------------------------------------------------

# VENN DIAGRAM: MICROBIOME PACKAGE (Nivel Género)

## v1 CORE MICROBIOME (GENUS) PLASTIC

Abundance >0.1%
Prevalence >= 50%

```{r}
library(microbiome)
plastic = subset_samples(physeq_R6_rel, SampleType2 == "Plastic")

# SELECT CORE TAXA AND CREATE A PHYLOSEQ OBJECT
pseq.core <- core(plastic, detection = 0.1, prevalence = .49)
# get the taxonomy data
tax.mat <- tax_table(pseq.core)
tax.df <- as.data.frame(tax.mat[,c(1:6)])
# Creat a string
tax.string = paste(tax.df$Rank1, tax.df$Rank2, tax.df$Rank3, tax.df$RAnk4, tax.df$Rank5, tax.df$Rank6, sep = ";")
```

Write this into a file:

```{r}
write.table(tax.string, "./Venn/Plastic-core-50.txt", sep = ",", dec = ".", row.names = FALSE, col.names = FALSE)
```

## v1 CORE MICROBIOME (GENUS) CAN

Abundance >0.1%
Prevalence >= 50%

```{r}
library(microbiome)
can = subset_samples(physeq_R6_rel, SampleType2 == "Can")

# SELECT CORE TAXA AND CREATE A PHYLOSEQ OBJECT
pseq.core <- core(can, detection = 0.1, prevalence = .49)
# get the taxonomy data
tax.mat <- tax_table(pseq.core)
tax.df <- as.data.frame(tax.mat[,c(1:6)])
# Creat a string
tax.string = paste(tax.df$Rank1, tax.df$Rank2, tax.df$Rank3, tax.df$RAnk4, tax.df$Rank5, tax.df$Rank6, sep = ";")
```

```{r}
write.table(tax.string, "./Venn/Can-core-50.txt", sep = ",", dec = ".", row.names = FALSE, col.names = FALSE)
```

## v1 CORE MICROBIOME (GENUS) CONTROL

Abundance >0.1%
Prevalence >= 50%

```{r}
library(microbiome)
control = subset_samples(physeq_R6_rel, SampleType2 == "Control")

# SELECT CORE TAXA AND CREATE A PHYLOSEQ OBJECT
pseq.core <- core(control, detection = 0.1, prevalence = .49)
# get the taxonomy data
tax.mat <- tax_table(pseq.core)
tax.df <- as.data.frame(tax.mat[,c(1:6)])
# Creat a string
tax.string = paste(tax.df$Rank1, tax.df$Rank2, tax.df$Rank3, tax.df$RAnk4, tax.df$Rank5, tax.df$Rank6, sep = ";")
```

```{r}
write.table(tax.string, "./Venn/Control-core-50.txt", sep = ",", dec = ".", row.names = FALSE, col.names = FALSE)
```

## v1 CORE MICROBIOME (GENUS) BIOFILM

Abundance >0.1%
Prevalence >= 50%

```{r}
library(microbiome)
biofilm = subset_samples(physeq_R6_rel, SampleType2 == "Biofilm")

# SELECT CORE TAXA AND CREATE A PHYLOSEQ OBJECT
pseq.core <- core(biofilm, detection = 0.1, prevalence = .49)
# get the taxonomy data
tax.mat <- tax_table(pseq.core)
tax.df <- as.data.frame(tax.mat[,c(1:6)])
# Creat a string
tax.string = paste(tax.df$Rank1, tax.df$Rank2, tax.df$Rank3, tax.df$RAnk4, tax.df$Rank5, tax.df$Rank6, sep = ";")
```

```{r}
write.table(tax.string, "./Venn/Biofilm-core-50.txt", sep = ",", dec = ".", row.names = FALSE, col.names = FALSE)
```

## Create the figure

The easiest way to create the Venn diagrams is to use this [web](http://bioinformatics.psb.ugent.be/webtools/Venn/) and the txt files as input.

-----------------------------------------------------------------------------------------------

# AMPVIS2 HEATMAP
## ALTERNATIVE TO BARPLOT

See [heatmaps con amvpis2](http://albertsenlab.org/ampvis2-heatmap/)

Load data into ampvis2:

```{r}
library(ampvis2)
if(!require("devtools"))
  install.packages("devtools")
#source the phyloseq_to_ampvis2() function from the gist
devtools::source_gist("8d0ca4206a66be7ff6d76fc4ab8e66c6")

# Necesitamos cambiar los colnames de la taxtable:
colnames(tax_table(physeq1_rel)) = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")
#go converty
ampvis2_obj <- phyloseq_to_ampvis2(physeq1_rel)

# Cambiamos el metadata para que la figura salga con el orden y los nombres que queremos
rownames(ampvis2_obj$metadata) = ampvis2_obj$metadata[,"SampleName"]
colnames(ampvis2_obj$abund) = ampvis2_obj$metadata[,"SampleName"]
```

As data is already normalised,  normalise = FALSE

## Heatmap Phylum

```{r}
p = amp_heatmap(
      data = ampvis2_obj,
      normalise = FALSE,
      tax_show = 20,
      tax_aggregate = "Phylum",
      plot_values_size = 3,
      min_abundance = 0.01,
      order_x_by = c("Control-1","Control-2","Control-3","Can-1", "Can-2", "Can-3", "Can-4","Plastic-1",
                     "Plastic-2","Plastic-3", "Plastic-4","Biofilm-1")
    )
p
# ggsave("./Amp_heatmap_Phylum.svg", p)
```

**OTHER VERSION**

```{r}
p = amp_heatmap(
      data = ampvis2_obj,
      normalise = FALSE,
      tax_show = 20,
      tax_aggregate = "Phylum",
      plot_values_size = 3,
      min_abundance = 0.01,
  color_vector = c("royalblue3",
                   "whitesmoke",
                   "lightcoral"),
      order_x_by = c("Control-1","Control-2","Control-3","Can-1", "Can-2", "Can-3", "Can-4","Plastic-1", "Plastic-2","Plastic-3", "Plastic-4","Biofilm-1")
    )
p
# ggsave("./Amp_heatmap_Phylum_v2.svg", p)
```

## Heatmap Genus

```{r}
library(ampvis2)
if(!require("devtools"))
  install.packages("devtools")
#source the phyloseq_to_ampvis2() function from the gist
devtools::source_gist("8d0ca4206a66be7ff6d76fc4ab8e66c6")

# Necesitamos cambiar los colnames de la taxtable:
colnames(tax_table(physeq_R6_rel)) = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")
#go converty
ampvis2_obj <- phyloseq_to_ampvis2(physeq_R6_rel)

# Cambiamos el metadata para que la figura salga con el orden y los nombres que queremos
rownames(ampvis2_obj$metadata) = ampvis2_obj$metadata[,"SampleName"]
colnames(ampvis2_obj$abund) = ampvis2_obj$metadata[,"SampleName"]
```


```{r}
p = amp_heatmap(
      data = ampvis2_obj,
      normalise = FALSE,
      tax_show = 20,
      tax_aggregate = "Genus",
      plot_values_size = 3,
      min_abundance = 0.01,
      color_vector = c("royalblue3",
                       "whitesmoke",
                       "lightcoral"),
      order_x_by = c("Control-1","Control-2","Control-3","Can-1", "Can-2", "Can-3", "Can-4","Plastic-1", "Plastic-2","Plastic-3", "Plastic-4","Biofilm-1")
    )
p
# ggsave("./Amp_heatmap_Genus.svg", p, limitsize = FALSE)
```


----------------------------------------------------------------

# CFU analysis

## Load the data

```{r}
CFU = read.csv2("./CFU-data/CFU-data.csv", sep = "\t")
head(CFU)
```

## Perform a wilcoxon test for all the samples

```{r}
res = c()

for (i in unique(CFU$Sample)) {
  aux = CFU[CFU$Sample == i,]
  plastic = aux[aux$CultureType == "Plastic", "CFU"]
  control = aux[aux$CultureType == "Control", "CFU"]
  res = c(res, wilcox.test(plastic, control)$p.value)
}

df = data.frame (unique(CFU$Sample), res)
colnames(df) = c("Sample", "p.value")
df

write.table(df, file = "wilcoxon-results.csv", sep = "\t", dec = ",",
            row.names = FALSE)
```

## Load the data for the figure

```{r}
CFU = read.csv2("./CFU-data/CFU-data-v2.csv", sep = "\t")
head(CFU)
```

Create the figure

```{r}
library(ggplot2)
library(RColorBrewer)

p = ggplot(data=CFU, aes(x=CultureType, y=CFUAvg, fill=CultureType)) + 
  facet_wrap( ~ Sample, scales = "free") +
  geom_bar(stat="identity", color="black", 
           position=position_dodge(), width = 0.8, alpha = 0.7) +
  geom_errorbar(aes(ymin=CFUAvg-SD, ymax=CFUAvg+SD), width=.2,
                 position=position_dodge(.9)) + 
  theme_light() + xlab("") + ylab("CFU") +
  theme(axis.text.x = element_blank()) +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(labels = function(x) format(x, scientific = TRUE))

p
```

--------------------------------------------------------------------------

# Accessory function (phyloseq_to_df)

```{r}
phyloseq_to_df <- function(physeq, addtax = T, addtot = F, addmaxrank = F, sorting = "abundance"){

  # require(phyloseq)

  ## Data validation
  if(any(addtax == TRUE || sorting == "taxonomy")){
    if(is.null(phyloseq::tax_table(physeq, errorIfNULL = F))){
      stop("Error: taxonomy table slot is empty in the input data.\n")
    }
  }

  ## Prepare data frame
  if(taxa_are_rows(physeq) == TRUE){
    res <- data.frame(OTU = phyloseq::taxa_names(physeq), phyloseq::otu_table(physeq), stringsAsFactors = F)
  } else {
    res <- data.frame(OTU = phyloseq::taxa_names(physeq), t(phyloseq::otu_table(physeq)), stringsAsFactors = F)
  }

  ## Check if the sample names were silently corrected in the data.frame
  if(any(!phyloseq::sample_names(physeq) %in% colnames(res)[-1])){
    if(addtax == FALSE){
      warning("Warning: Sample names were converted to the syntactically valid column names in data.frame. See 'make.names'.\n")
    }

    if(addtax == TRUE){
      stop("Error: Sample names in 'physeq' could not be automatically converted to the syntactically valid column names in data.frame (see 'make.names'). Consider renaming with 'sample_names'.\n")
    }
  }

  ## Add taxonomy
  if(addtax == TRUE){

    ## Extract taxonomy table
    taxx <- as.data.frame(phyloseq::tax_table(physeq), stringsAsFactors = F)

    ## Reorder taxonomy table
    taxx <- taxx[match(x = res$OTU, table = rownames(taxx)), ]

    ## Add taxonomy table to the data
    res <- cbind(res, taxx)

    ## Add max tax rank column
    if(addmaxrank == TRUE){

      ## Determine the lowest level of taxonomic classification
      res$LowestTaxRank <- get_max_taxonomic_rank(taxx, return_rank_only = TRUE)

      ## Reorder columns (OTU name - Taxonomy - Max Rank - Sample Abundance)
      res <- res[, c("OTU", phyloseq::rank_names(physeq), "LowestTaxRank", phyloseq::sample_names(physeq))]

    } else {
      ## Reorder columns (OTU name - Taxonomy - Sample Abundance)
      res <- res[, c("OTU", phyloseq::rank_names(physeq), phyloseq::sample_names(physeq))]

    } # end of addmaxrank
  }   # end of addtax

  ## Reorder OTUs
  if(!is.null(sorting)){

    ## Sort by OTU abundance
    if(sorting == "abundance"){
      otus <- res[, which(colnames(res) %in% phyloseq::sample_names(physeq))]
      res <- res[order(rowSums(otus, na.rm = T), decreasing = T), ]
    }

    ## Sort by OTU taxonomy
    if(sorting == "taxonomy"){
      taxtbl <- as.data.frame( phyloseq::tax_table(physeq), stringsAsFactors = F )

      ## Reorder by all columns
      taxtbl <- taxtbl[do.call(order, taxtbl), ]
      # taxtbl <- data.table::setorderv(taxtbl, cols = colnames(taxtbl), na.last = T)
      res <- res[match(x = rownames(taxtbl), table = res$OTU), ]
    }
  }

  ## Add OTU total abundance
  if(addtot == TRUE){
    res$Total <- rowSums(res[, which(colnames(res) %in% phyloseq::sample_names(physeq))])
  }

  rownames(res) <- NULL
  return(res)
}
```

